Parallel global optimization algorithm employing decision trees for launching local methods

### 2025-10-14 2025/MMST_DecisionTree/ParallelDecisionTree.tex
Add description of Globalizer

In each iteration, the algorithm selects several computation points and distributes them to processes in groups. MPI manages this, with process zero handling point distribution and result collection, but remaining computationally idle. Each remaining process receives a group of points, which are computed in parallel using the CPU.

@@

	\item Compute the coordinates of the new trial points $y'^{k+j}=y\left(x^{k+j}\right), \ 1\leq j\leq n$, whose preimages are  $x^{k+j}\in\left(x_{t_j-1},x_{t_j}\right)$, according to the following formula:
	\begin{equation}
		\label{agp5_x1}
		x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}\left(z_{t_j}-z_{t_j-1}\right)\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{M}\right]^N.
	\end{equation}

	\item Send to process $j$, $1 \leq j \leq n$, the coordinates of points $y'^{k+1}, \dots, y'^{k+n}$ and notify the process that it needs to perform a single trial.

@@

	\item Add the trial points $y'^{k+1}=y\left(x^{k+1}\right) , \dots, y'^{k+n}=y\left(x^{k+n}\right)$  to the set $V$.

@@

	\item 
	%If the decision tree is being used for the first time, construct a uniform grid
	Upon the first application of the decision tree, generate a uniform grid
	\begin{displaymath} 
		Y'=\{ y'\in R^N:\ a_i\le y'_i \le b_i,\ 1\le i\le N \},
	\end{displaymath} 
	%where the number of grid nodes is a parameter of the method.
	where the density of grid points is governed by a user-defined parameter.

@@

	\item For all points $y' \in V$, find the points $y'_q$ closest to $y'$.
		If no point $y'_q$ has a value smaller than $y'$, add $y'$ to the set $U$.
		Clear the set $V$.

@@

>>>>

In each iteration, the algorithm selects several computation points and distributes them to processes in groups. MPI (Message Passing Interface) manages this, with process zero handling point distribution and result collection, but remaining computationally idle. Each remaining process receives a group of points, which are computed in parallel using the CPU.

@@

	\item Compute the new trial points $\overline{y}^{k+j}=y\left(x^{k+j}\right), \ 1\leq j\leq n$, whose preimages $x^{k+j}\in\left(x_{t_j-1},x_{t_j}\right)$  are calculated using the following formula:
	\begin{equation}
		\label{agp5_x1}
		x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}\left(z_{t_j}-z_{t_j-1}\right)\frac{1}{2r}\left[\frac{\left|z_{t_j}-z_{t_j-1}\right|}{M}\right]^N.
	\end{equation}

	\item Send to process $j$, $1 \leq j \leq n$, the coordinates of points $\overline{y}^{k+1}, \dots, \overline{y}^{k+n}$ and notify the process that it needs to perform a single trial.

@@

	\item Add the trial points $\overline{y}^{k+1}=y\left(x^{k+1}\right) , \dots, \overline{y}^{k+n}=y\left(x^{k+n}\right)$  to the set $V$.

@@

	\item 
	%If the decision tree is being used for the first time, construct a uniform grid
	Upon the first application of the decision tree, generate a uniform grid
	\begin{displaymath} 
		\overline{Y}=\{ \overline{y}\in R^N:\ a_i\le \overline{y}_i \le b_i,\ 1\le i\le N \},
	\end{displaymath} 
	%where the number of grid nodes is a parameter of the method.
	where the density of grid points is governed by a user-defined parameter.

@@

	\item For all points $\overline{y} \in V$, find the points $\overline{y}_q$ closest to $\overline{y}$.
		If no point $\overline{y}_q$ has a value smaller than $\overline{y}$, add $\overline{y}$ to the set $U$.
		Clear the set $V$.

@@

The software implementation of the described algorithm is an extension of the Globalizer framework (https://github.com/OptimLLab/Globalizer). The development was carried out in C++ using the OpenCV library and MPI.

The Globalizer framework is designed for solving multiextremal optimization problems on supercomputer systems. Typical areas of application of the software package include solving problems of identifying parameters of mathematical models based on experimental data, as well as optimizing parameters of artificial intelligence models and other complex systems.
The framework is characterized by the following features:
\begin{itemize}
    \item operates on Windows and Linux family operating systems;
    \item provides a standard interface in this subject area;
    \item provides the ability to work using default algorithm parameter values that ensure a high-quality solution;
    \item supports parallelism using MPI and OpenMP technologies;
    \item supports the use of computing accelerators.
\end{itemize}
